A thermodynamic view of dusty protoplanetary disks

### 2017-07-13 paper.tex
minor tweaks to intro which I'm happy with.  Remaining concern is whether we should clarify what general vs. explicit stability analyses are (i.e. for a specific disk model or not).

The advantage of this reformulation arises in the tight coupling limit of  $\tstop \ll 1/\Omega$, where the orbital frequency $\Omega$ sets the characteristic timescale in many disk dynamics problems.
Models with two fluids or with Lagrangian solids are numerically difficult to evaluate in this regime, due to stiff drag forces.  
In the reformulated equations, however, the center of mass motion does
not experience drag forces and thus is not stiff.  Moreover,
for $\tstop\Omega\ll 1$ the relative motion satisfies the terminal-velocity approximation and can be eliminated from the equations.  This single fluid framework for a well-coupled dust-gas mixture facilitates numerical calculations \citep{price15}; and also simplifies analytic calculations \citep{youdin05a,jacquet11}, which is useful for gaining insight.

@@

In this work, we develop a parallel between this single-fluid description of dust-gas mixtures and standard hydrodynamics.
Specifically we recast the evolution of the dust-to-gas ratio as an energy equation for the thermal content of a fluid.
This parallel is most precise when the gas in the dust/gas mixture obeys a locally isothermal equation of state for the gas, $P = c_s^2\rhog$, where $c_s$ is a prescribed sound-speed profile.  Thus we make this approximation which holds when cooling times are short \citep{lin15}, as with protoplanetary disks 
 whose temperature is set by external irradiation
\citep{chiang97,stam08}.

@@

This paper is organized as follows. 
%AY: or not? %We first present, without
%derivation, the correspondence between dusty gas and     
%ideal hydrodynamics in \S\ref{key_idea}. 
In \S\ref{grow_osc} 
we describe a key insight of our study, that pressure (i.e. ``PdV") work explains the basis of several instabilities, including SI and VSI, in terms of
pressure-density phase lags.   
%This means that dusty-gas phenomena, such 
%as SI and VSI, have thermodynamical interpretations.  
We begin our formal discussion in \S\ref{setup} by transforming the
basic equations describing tightly-coupled dusty gas into
hydrodynamic equations with a special cooling function.  Here we also
formalize the concept of dust-induced entropy and buoyancy. 
%We derive equilibrium solutions in \S\ref{eqm}, which we use in our
%explicit stability calculations.
In \S\ref{limits} we analyze the general stability properties of dusty
disks. We then move onto explicit stability analyses in
\S\ref{linear_problem}, were we show that dusty edges are unstable,
and revisit the SGI and SI in the tight
coupling regime. In \S\ref{results} we
extend the VSI to dusty disks. We   
discuss future applications of the dusty/adiabatic gas equivalence in 
\S\ref{discussion} before summarizing in \S\ref{summary}.

>>>>

This reformulation is particularly advantageous in the tight coupling limit of  $\tstop \ll 1/\Omega$, where the orbital frequency $\Omega$ sets the characteristic timescale in many disk dynamics problems.
Models with two fluids or with Lagrangian solids are numerically difficult to evaluate in this regime, due to stiff drag forces.  
In the reformulated equations, however, the center of mass motion does
not experience drag forces and thus is not stiff.  Moreover,
for $\tstop\Omega\ll 1$ the relative motion satisfies the terminal-velocity approximation and can be eliminated from the equations.  This single fluid framework for a well-coupled dust-gas mixture facilitates numerical calculations \citep{price15}; and also simplifies analytic calculations \citep{youdin05a,jacquet11}, which is useful for gaining insight.

@@

In this work, we develop a parallel between this single-fluid description of dust-gas mixtures and standard hydrodynamics.
Specifically we recast the evolution of the dust-to-gas ratio as an energy equation for the thermal content of a fluid.
This parallel is most precise when the gas in the dust/gas mixture obeys a locally isothermal equation of state for the gas, $P = c_s^2\rhog$, where $c_s$ is a prescribed sound-speed profile.   This isothermal approximation holds when cooling times are short \citep{lin15}, e.g.\ in protoplanetary disks 
 whose temperature is set by external irradiation
\citep{chiang97,stam08}.

@@

This paper is organized as follows. 
In \S\ref{grow_osc} 
we describe a key insight of our study, that pressure (i.e.\ ``PdV") work explains the basis of several instabilities, including SI and VSI, in terms of
pressure-density phase lags.    
We develop our formalism in \S\ref{setup} by transforming the
two fluid equations of a tightly-coupled dusty gas into single fluid
hydrodynamic equations with a special cooling function.  Here we also
define the effective entropy and buoyancy of our model fluid. 
%We derive equilibrium solutions in \S\ref{eqm}, which we use in our
%explicit stability calculations.
In \S\ref{limits} we analyze the general stability properties of dusty
disks. We then move onto explicit stability analyses in
\S\ref{linear_problem}, were we show that dusty edges are unstable,
and revisit the SGI and SI in the tight
coupling regime. In \S\ref{results} we
extend the VSI to dusty disks. We   
discuss future applications of the dusty/adiabatic gas equivalence in 
\S\ref{discussion} before summarizing in \S\ref{summary}.
